
---
title: 'Fitbit Feature Engineering 2 : Predict sleep_score without stress_score'
output: pdf_document
---

GOAL : predict sleep_score without stress_score\\
CONCLUSION : Model m4 has the highest rsq and lowest mrse. m4 is a linear regression model that uses predictors date and deep_sleep_min only.

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(tidymodels)
library(car)
library(yardstick)
```

```{r}
fitbit_df <- read.csv('fitbit_data.csv')
fitbit_df <- fitbit_df %>% select(-stress_score)
fitbit_df$date <- as.Date(fitbit_df$date)
head(fitbit_df)
```

# Split Data
```{r}
set.seed(123)
split <- initial_split(fitbit_df, prop=0.9)
train <- training(split)
test <- testing(split)
```

# Influential Points
```{r}
lm <- lm(data=fitbit_df, sleep_score ~ .)
summary(lm)
par(mfrow=c(2,2))
plot(lm)
```

## Outliers : 36 122 124 132 133
```{r}
q <- quantile(train$sleep_score, c(0.25, 0.75))
iqr <- IQR(train$sleep_score)
threshold <- 1.5 * iqr
outliers <- which(train$sleep_score < (q[1] - threshold) | train$sleep_score > (q[2] + threshold))
print(outliers)

boxplot(train$sleep_score, main = "Boxplot of Sleep Score with Outliers",
        ylab = "Sleep Score", col = "lightblue", pch = 19)
points(outliers, train$sleep_score[outliers], col = "red", pch = 19)
```
# New train set after removing outliers
```{r}
train <- train[-outliers]
```



# m1 : all predictors
```{r}
m1 <- linear_reg()

m1_recipe <- recipe(data=train, sleep_score ~ .) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_date(date, features = c("dow", "month", "year")) %>%
  step_holiday(date) %>%
  step_corr(all_numeric_predictors(), threshold = 0.5) %>%
  step_YeoJohnson(all_numeric_predictors()) 
  

m1_wkfl <- workflow() %>%
  add_model(m1) %>%
  add_recipe(m1_recipe)

m1_fit <- m1_wkfl %>%
  fit(data=train)

m1_aug <- m1_fit %>% 
  augment(test) 

m1_aug %>% 
  metrics(truth = sleep_score, estimate = .pred)
```

# m2 : 5 selected predictors based on my expectation
```{r}
m2 <- linear_reg()

m2_recipe <- recipe(data=train, sleep_score ~ date+deep_sleep_min+AZM_minutes+o2_avg+resting_heart_rate) %>%
    step_normalize(all_numeric_predictors()) %>%
    step_date(date, features = c("dow", "month", "year")) %>%
    step_holiday(date) %>%
    step_corr(all_numeric_predictors(), threshold = 0.5) %>%
    step_YeoJohnson(all_numeric_predictors()) 
  
m2_wkfl <- workflow() %>%
  add_model(m2) %>%
  add_recipe(m2_recipe)
  
m2_fit <- m2_wkfl %>%
  fit(data=train)
  
m2_aug <- m2_fit %>% 
  augment(test) 
  
m2_aug %>% 
  metrics(truth = sleep_score, estimate = .pred)
```

# m3 : 3 selected predictors based on importance
```{r}
library(vip)
m1_fit %>%
  extract_fit_parsnip() %>%
  vip(aesthetics = list(fill = "#00B0B9"))
```

```{r}
m3 <- linear_reg()

m3_recipe <- recipe(data=train, sleep_score ~ date+deep_sleep_min+AZM_minutes) %>%
    step_normalize(all_numeric_predictors()) %>%
    step_date(date, features = c("dow", "month", "year")) %>%
    step_holiday(date) %>%
    step_corr(all_numeric_predictors(), threshold = 0.5) %>%
    step_YeoJohnson(all_numeric_predictors()) 

m3_wkfl <- workflow() %>%
  add_model(m3) %>%
  add_recipe(m3_recipe)

m3_fit <- m3_wkfl %>%
  fit(data=train)

m3_aug <- m3_fit %>%
  augment(test)

m3_aug %>% 
  metrics(truth = sleep_score, estimate = .pred)
```

# m4 : 2 selected predictors
rsq = 0.8137866	
Highest rsq : after k=2 and remove influential -> 0.8291957	
```{r}
m4 <- linear_reg()

m4_recipe <- recipe(data=train, sleep_score ~ date+deep_sleep_min) %>%
    step_normalize(all_numeric_predictors()) %>%
    step_date(date, features = c("dow", "month", "year")) %>%
    step_holiday(date) %>%
    step_corr(all_numeric_predictors(), threshold = 0.5) %>%
    step_YeoJohnson(all_numeric_predictors()) 

m4_wkfl <- workflow() %>%
  add_model(m4) %>%
  add_recipe(m4_recipe)

m4_fit <- m4_wkfl %>%
  fit(data=train)

m4_aug <- m4_fit %>%
  augment(test)

m4_aug %>% 
  metrics(truth = sleep_score, estimate = .pred)
```
